**Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA sequencing system**

### 2016-01-30 Schloss_PacBio16S_PeerJ_2016.Rmd
Extract taxonomy analysis

```{r error_profile}
error_profile <- fromJSON(scan("data/process/error_profile.json", quiet=T, what="character"))

error_summary <- read.table(file="data/process/error_summary.tsv", header=T)
raw_to_basic_lost <- 100 * (1 - error_summary[error_summary$method == "basic","nseqs"]/error_summary[error_summary$method == "raw","nseqs"])

@@

@@

```{r random_stuff }

@@

```{r, classificationSummary}
data <- read.table(file="data/process/taxonomy.depth.analysis", header=T)

rdp <- 100 * cbind("Mock"=data[data$db=="pds" & data$dataset=="mock", "genus_species"],
				 "Human"=data[data$db=="pds" & data$dataset=="human", "genus_species"],
				 "Mouse"=data[data$db=="pds" & data$dataset=="mouse", "genus_species"],
				 "Soil"=data[data$db=="pds" & data$dataset=="soil", "genus_species"])
rownames(rdp) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")

gg.sp <- 100 * cbind("Mock"=data[data$db=="gg" & data$dataset=="mock", "X7"],
					 "Human"=data[data$db=="gg" & data$dataset=="human", "X7"],
					 "Mouse"=data[data$db=="gg" & data$dataset=="mouse", "X7"],
					 "Soil"=data[data$db=="gg" & data$dataset=="soil", "X7"])
rownames(gg.sp) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")
```

***Increasing sequence length improves classification.***
We classified all of the sequence data we generated using the naïve Bayesian
classifier using the RDP, SILVA, and greengenes reference taxonomies (Figure
4). In general, increasing the length of the region improved the
ability to assign the sequence to a genus or species. Interestingly, each of the
samples we analyzed varied in the ability to assign its sequences to the depth of
genus or species. Furthermore, the reference database that did the best job of
classifying the sequences varied by sample type. For example, the SILVA
reference did the best for the human feces and soil samples and the RDP did the
best for the mouse feces samples. An advantage of the greengenes database is
that it contains
information for 2,514 species-level lineages for 11% of the reference sequences;
the other databases only provided taxonomic data to the genus level. There was a
modest association between the length of the fragment and the
ability to classify sequences to the species-level for the human samples; there
was no such association for the mouse and soil samples. In fact, at most
`r format(round(max(gg.sp[,"Soil"]), digits=1), nsmall=1)`% of
the soil sequences and `r format(round(max(gg.sp[,"Mouse"]), digits=1), nsmall=1)`%
of the mouse sequences could be classified to a
species. These results indicate that the ability to classify sequences to the
genus or species level is a function of read length, sample type, and the
reference database.

@@

[ This figure removed ]
**Figure 3. The relationship between the error rate of each region and the
composite quality scores for the sequences.** The error rates (A and C) and
percentage of sequences (B and D) were calculated for the reads that had a
composite quality score above the plotted value. The composite quality scores
were calculated by either determining the minimum value of the average quality
score within all 50-nt windows within each region (A and B) or by calculating
the average quality score across the entire sequence read (C and D).

**Figure 4. Percentage of unique sequences that could be classified.**
Classifications were performed using taxonomy references curated from the RDP,
SILVA, or greengenes databases for the four types of samples that were sequenced
across the six regions from the 16S rRNA gene. Only the greengenes taxonomy
reference provided species-level information.

@@

### Figure 2
```{r Figure_2, width=6.5, height=6}
error.fraction <- cbind(initial.error=init.rates, basic.error=basic.error,
							bcprimer.error, coverage.error, aveq.error,
							oligosCoverage.error, oligosQScore.error,
							coverageQScore.error, allFilters.error, final.error)
rownames(error.fraction) <- gsub("v(\\d*)", "\\1", rownames(error.fraction))

@@

### Figure 3
```{r Figure_3, fig.height=6.5, fig.width=6.5}
aggregateErrorMinQualityScore <- function(folder){
	write(folder, "")
	composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder,
								 ".composite", sep=""), header=T, row.names=1)
	composite <- composite[composite$reason == "g",]

	remaining.error <- rep(0, max(composite$minQ)+1)
	qscores <- 0:max(composite$minQ)

	remaining.error <- unlist(lapply(qscores, function(x){mean(composite[composite$minQ >= x,"error"])}))
	names(remaining.error) <- qscores
	points(x=10:70, remaining.error[10:70], type="l", col=clrs[folder], lwd=2)
}

fractionMinQualityScore <- function(folder){
	write(folder, "")

	composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder,
					 				".composite", sep=""), header=T, row.names=1)
	composite <- composite[composite$reason == "g",]

	binned <- aggregate(composite$error, by=list(round(composite$minQ)), length)
	remaining <- 1-cumsum(binned$x)/sum(binned$x)
	points(binned$Group.1, remaining, type="l", col=clrs[folder], lwd=2)
}

aggregateErrorAveQualityScore <- function(folder){
	write(folder, "")
	composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder,
								 ".composite", sep=""), header=T, row.names=1)
	composite <- composite[composite$reason == "g",]

	remaining.error <- rep(0, max(composite$aveQ)+1)
	qscores <- 0:max(composite$aveQ)
	remaining.error <- unlist(lapply(qscores,
					function(x){mean(composite[composite$aveQ >= x,"error"])}))
	names(remaining.error) <- qscores
	points(x=10:70, remaining.error[10:70], type="l", col=clrs[folder], lwd=2)
}

fractionAveQualityScore <- function(folder){
write(folder, "")

	composite <- read.table(file=paste("pipeline_dev/", folder, "/", folder,
								 ".composite", sep=""), header=T, row.names=1)
	composite <- composite[composite$reason == "g",]

	binned <- aggregate(composite$error, by=list(round(composite$aveQ)), length)

	remaining <- 1-cumsum(binned$x)/sum(binned$x)
	points(binned$Group.1, remaining, type="l", col=clrs[folder], lwd=2)
}

layout(1)
layout(matrix(c(1,2,3,4), nrow=2, byrow=F), width=c(1.1,1), height=c(1, 1.3))

#A
par(mar=c(0.5,5,0.5, 0.5))
plot(1, xlim=c(10,75), ylim=c(0,0.015), type="n", axes=F, xlab="", ylab="")
lapply(regions, aggregateErrorMinQualityScore)
mtext(side=2, line=2.5, text="Error rate of sequences\nabove threshold (%)", cex.lab=0.9)
axis(1, at=seq(10,70,10), label=F, tick=T)
axis(2, las=2, at=seq(0,0.015,0.005), label=format(seq(0,1.5,0.5), nsmall=1), cex.axis=1.25)
box()
mtext(side=2, line=2.6, at=0.015, text="A", las=2, cex=2, font=2)

#B
par(mar=c(6,5,0.5, 0.5))
plot(1, xlim=c(10,75), ylim=c(0,1), type="n", yaxt="n", xlab="", ylab="", cex.axis=1.25)
lapply(regions, fractionMinQualityScore)
mtext(side=1, line=4.25, text="Minimum average quality\nscore within a 50-nt window\nacross the full sequence", cex=0.9)
mtext(side=2, line=2.5, text="Sequences above\nthreshold (%)", cex.lab=0.9)

axis(2, las=2, at=seq(0,1,0.2), label=seq(0,100,20), cex.axis=1.25)
mtext(side=2, line=2.6, at=1, text="B", las=2, cex=2, font=2)

#C
par(mar=c(0.5,2,0.5, 0.5))
plot(1, xlim=c(10,75), ylim=c(0,0.015), type="n", axes=F, xlab="", ylab="")
lapply(regions, aggregateErrorAveQualityScore)
axis(1, at=seq(10,70,10), label=F, tick=T)
axis(2, at=seq(0,0.02,0.005), label=F, tick=T)
box()
mtext(side=2, line=0.6, at=0.015, text="C", las=2, cex=2, font=2)

#D
par(mar=c(6,2,0.5, 0.5))
plot(1, xlim=c(10,75), ylim=c(0,1), type="n", yaxt="n", xlab="", ylab="", cex.axis=1.25)
lapply(regions, fractionAveQualityScore)
mtext(side=1, text="Average quality score", cex=0.9, line=3)
title(ylab="Sequences above threshold (%)")
axis(2, at=seq(0,1,0.2), label=FALSE, tick=TRUE)
mtext(side=2, line=0.6, at=1, text="D", las=2, cex=2, font=2)

legend(x=10, y=0.50, legend=pretty_region[regions], lty=1, lwd=2, col=clrs[regions], cex=1.0)
```

### Figure 4

```{r, Figure_4, fig.width=7, fig.height=7}
par(mar=c(5, 5, 0.5, 1))
dotchart(gg, xlim=c(0,100), col="black", xlab="Unique reads that classified\nto genus or species level (%)", pch=19)

>>>>

```{r error_profile}
error_profile <- fromJSON(scan("data/process/error_profile.json", quiet=T, what="character"))
```

@@

```{r basic}
error_summary <- read.table(file="data/process/error_summary.tsv", header=T)

raw_to_basic_lost <- 100 * (1 - error_summary[error_summary$method == "basic","nseqs"]/error_summary[error_summary$method == "raw","nseqs"])
```

@@

@@

```{r, classificationSummary}
data <- read.table(file="data/process/taxonomy_depth_analysis.tsv", header=T)
```

***Increasing sequence length improves classification.***
We classified all of the sequence data we generated using the naïve Bayesian classifier using the RDP, SILVA, and greengenes reference taxonomies (Figure 3). In general, increasing the length of the region improved the ability to assign the sequence to a genus or species. Interestingly, each of the samples we analyzed varied in the ability to assign its sequences to the depth of genus or species. Furthermore, the reference database that did the best job of classifying the sequences varied by sample type. For example, the SILVA reference did the best for the human feces and soil samples and the RDP did the best for the mouse feces samples. An advantage of the greengenes database is that it contains information for 2,514 species-level lineages for 11% of the reference sequences; the other databases only provided taxonomic data to the genus level. There was a modest association between the length of the fragment and the ability to classify sequences to the species-level for the human samples; there was no such association for the mouse and soil samples. In fact, at most `r format(round(100*max(data[data$db=="gg" & data$dataset=="soil","X7"]), digits=1), nsmall=1)`% of the soil sequences and `r format(round(100*max(data[data$db=="gg" & data$dataset=="mouse","X7"]), digits=1), nsmall=1)`% of the mouse sequences could be classified to a species. These results indicate that the ability to classify sequences to the genus or species level is a function of read length, sample type, and the reference database.

@@

**Figure 3. Percentage of unique sequences that could be classified.**
Classifications were performed using taxonomy references curated from the RDP,
SILVA, or greengenes databases for the four types of samples that were sequenced
across the six regions from the 16S rRNA gene. Only the greengenes taxonomy
reference provided species-level information.

@@

### Figure 2
```{r random_stuff }

```{r Figure_2, width=6.5, height=6}
error.fraction <- cbind(initial.error=init.rates, basic.error=basic.error,
							bcprimer.error, coverage.error, aveq.error,
							oligosCoverage.error, oligosQScore.error,
							coverageQScore.error, allFilters.error, final.error)
rownames(error.fraction) <- gsub("v(\\d*)", "\\1", rownames(error.fraction))

@@

### Figure 3

```{r, Figure_4, fig.width=7, fig.height=7}
rdp <- 100 * cbind("Mock"=data[data$db=="pds" & data$dataset=="mock", "genus_species"],
				 "Human"=data[data$db=="pds" & data$dataset=="human", "genus_species"],
				 "Mouse"=data[data$db=="pds" & data$dataset=="mouse", "genus_species"],
				 "Soil"=data[data$db=="pds" & data$dataset=="soil", "genus_species"])
rownames(rdp) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")

gg.sp <- 100 * cbind("Mock"=data[data$db=="gg" & data$dataset=="mock", "X7"],
					 "Human"=data[data$db=="gg" & data$dataset=="human", "X7"],
					 "Mouse"=data[data$db=="gg" & data$dataset=="mouse", "X7"],
					 "Soil"=data[data$db=="gg" & data$dataset=="soil", "X7"])
rownames(gg.sp) <- c("V4", "V3-V5", "V1-V3", "V1-V5", "V1-V6", "V1-V9")

par(mar=c(5, 5, 0.5, 1))
dotchart(gg, xlim=c(0,100), col="black", xlab="Unique reads that classified\nto genus or species level (%)", pch=19)
